---
title: "W1-1_SettingUpAndExploringTheData"
author: "Taufiqur Rohman"
date: '2022-03-31'
output: html_document
---

Read the data using read.delim, because the txt source file is separated by tab.

```{r read data frame}
df_purchase <- read.delim("purchases.txt", header = FALSE, sep = "\t", dec = ".")
```

Display what has loaded in the data

```{r explore data}
head(df_purchase) # Check first 5 first records
summary(df_purchase) # Check mean, median, max, min, etc.
str(df_purchase) # Check the data type
```

I found that there are no column names in the data. So, I label all of the columns using the colnames() function.

```{r adding column names}
colnames(df_purchase) <- c("cust_id", "purchase_amt", "date_of_purchase")
```

I also found that the the data format of "date_of_purchase" is still numeric. Let's convert it to date.

```{r converting to date format}
df_purchase$date_of_purchase <- as.Date(df_purchase$date_of_purchase, "%Y-%m-%d")
df_purchase$year_of_purchase <- as.numeric(format(df_purchase$date_of_purchase, "%Y"))
```

Let's check it out again the end to confirm the data set is clean and ready to be analyzed.

```{r explore data again}
head(df_purchase)
summary(df_purchase)
str(df_purchase)
```

In the course, they said me to explore the database using the SQL statements. So, let's explore the summary of the data using the SQL statements.

```{r install sqldf}
library(sqldf)
```

```{r sql statements}
sql_exp_1 <- sqldf("SELECT 
                      year_of_purchase, 
                      COUNT(year_of_purchase) AS 'counter',
                      AVG(purchase_amt) AS 'avg_purchase',
                      SUM(purchase_amt) AS 'total_purchase'
                    FROM 
                      df_purchase
                    GROUP BY
                      1
                    ORDER BY
                      1")
```

In the end, let's visualize them into a bar plot to make it easier to understand by the audience.

```{r visualization}
# How many purchases over year
barplot(sql_exp_1$counter, names.arg = sql_exp_1$year_of_purchase)

# Average purchases amount every year
barplot(sql_exp_1$avg_purchase, names.arg = sql_exp_1$year_of_purchase)

# Total amount of purchases every year
barplot(sql_exp_1$counter, names.arg = sql_exp_1$year_of_purchase)
```

