---
title: "W2-2_QuizModule1"
author: "Taufiqur Rohman"
date: '2022-04-01'
output: html_document
---

## Reading and Cleaning the Data

Read the data using read.delim, because the txt source file is separated by tab.

```{r read data frame}
df_purchase <- read.delim("purchases.txt", header = FALSE, sep = "\t", dec = ".")
```

Display what has loaded in the data

```{r explore data}
head(df_purchase) # Check first 5 first records
summary(df_purchase) # Check mean, median, max, min, etc.
str(df_purchase) # Check the data type
```

I found that there are no column names in the data. So, I label all of the columns using the colnames() function.

```{r adding column names}
colnames(df_purchase) <- c("cust_id", "purchase_amt", "date_of_purchase")
```

I also found that the the data format of "date_of_purchase" is still numeric. Let's convert it to date. Also, I want to know the how many days elapsed to the date of purchase since January 1st, 2016. This will be helpful for us to calculate the Recency.

```{r converting to date format}
df_purchase$date_of_purchase <- as.Date(df_purchase$date_of_purchase, "%Y-%m-%d")
df_purchase$days_since <- as.numeric(difftime(time1 = "2016-01-01",
                                           time2 = df_purchase$date_of_purchase,
                                           units = "days"))
```

Let's check it out again the end to confirm the data set is clean and ready to be analyzed.

```{r explore data again}
head(df_purchase)
summary(df_purchase)
str(df_purchase)
```

For now, lets calculate the RFM. I will use the sqldf package so I can write SQL statement in R environment as it is easier to do that.

```{r sql statement}
sql_rfm <- sqldf("SELECT
                    cust_id,
                    MIN(days_since) as 'R_recency',
                    COUNT(*) as 'F_frequency',
                    AVG(purchase_amt) as 'M_monval'
                  FROM
                    df_purchase
                  GROUP BY
                    cust_id")
```

## Visualizing the Data

To see the data distribution clearer, I would like to create several visualizations to help us in understanding the data.

```{r using histogram to read data}
# Big picture of the data
summary(sql_rfm) 

# Distribution of the data
hist(sql_rfm$R_recency, main = "Recency Distribution", xlab = "Days Elapsed")
hist(sql_rfm$F_frequency, main = "Frequency Distribution", xlab = "Number of Purchases")
hist(sql_rfm$M_monval, main = "Monetary Value Distribution", xlab = "in USD", breaks = 100)
```

## Preparing Data Frame for Segmentation

The data is clean and set. Now, for the last thing, as we want to do the segmentation analysis, let's prepare so they are ready to be analyzed.

```{r}
# Duplicate the data frame
df_purchase_new <- sql_rfm

# Make Customer ID as the row name
row.names(df_purchase_new) <- df_purchase_new$cust_id
df_purchase_new$cust_id <- NULL
head(df_purchase_new)

# Transform monetary value and frequency variable to log as it is very skewed to the left
df_purchase_new$M_monval <- log(df_purchase_new$M_monval)
df_purchase_new$F_frequency <- log(df_purchase_new$F_frequency)
hist(df_purchase_new$M_monval, main = "Log Monetary Value Distribution", xlab = "in USD")
hist(df_purchase_new$F_frequency, main = "Log Frequency Distribution", xlab = "Number of Purchase")

# Standardizing all of the variables
df_purchase_new <- scale(df_purchase_new)

# Final check
head(df_purchase_new)
summary(df_purchase_new)
```

## Running the Hierarchical Segmentation

The data is ready. I will perform the hierarchical segmentation now.

```{r preparing the data set}
## Split train set using 10% of the sample size
smp <- seq(1, 18417, by = 10)
head(smp)

df_sample <- sql_rfm[smp, ]
df_train <- df_purchase_new[smp, ]
```

```{r performing clustering}
# Performing clustering
d <- dist(df_train)
c <- hclust(d, method = "ward.D2")

plot(c, xlab = "Customers")

# Cutting to the optimum cluster
# Based on statistical fit (dendogram) and managerial relevance
segments <- cutree(c, k = 5)
```



```{r exploring the data} 
segments[1:30]
table(segments)
```

```{r checking the aggregate of every segments}
segment_final <- aggregate(df_sample[ , 2:4], by = list(segments), FUN = mean)
```

By doing this analysis, I have successfully created 9 groups for this customer data based on their Recency, Frequency, and amount of money spent in the company.

## QUESTION

1. What is the size of the largest segment?

```{r question 1}
df_sample$segment <- segments
table(df_sample$segment)
```

2. What is the average purchase amount of the segment which contains, on average, the customers who have made their last purchase the most recently?

```{r question 2}
segment_final
```

3. Of the five following criteria to determine the best number of segments, which one is the least relevant?

```{r question 3}
```

4. Does customer #260 belongs to the same segment than customer #5920?

```{r question 4}
df_sample %>%
  filter(cust_id == "260") %>% 
  select(segment)
df_sample %>%
  filter(cust_id == "5920") %>% 
  select(segment)
```

5. Looking at the average profile of segment 1, would you say that members of this segment are typically... (you'll have to choose the right proposition)

```{r question 5}
segment_final
```
