---
title: "W3-2_QuizModule2"
author: "Taufiqur Rohman"
date: '2022-04-04'
output: html_document
---

To validate Module 2 and correctly answer the questions will require that you perform the following exercise first: take the code in the file named module2.R, and modify it such that the managerial segment "new active" is divided into two sub-segments called "new active high" (with an average purchase amount of 100 dollar or more) and "new active low" (with an average purchase amount of less than 100 dollar). Tips: make sure that you apply that modification to both the 2015 and 2014 segmentations, and that you modify accordingly the code that re-order the factor "segment".

So, I will modify the coding that I have performed in the previous exercise (refer to W3-1_CodingManagerialSegmentation)

## Reading and Cleaning the Data

```{r read data frame}
df_purchase <- read.delim("purchases.txt", header = FALSE, sep = "\t", dec = ".")
colnames(df_purchase) <- c("cust_id", "purchase_amt", "date_of_purchase")

# Converting data format and adding Recency column
df_purchase$date_of_purchase <- as.Date(df_purchase$date_of_purchase, "%Y-%m-%d")
df_purchase$days_since <- as.numeric(difftime(time1 = "2016-01-01",
                                           time2 = df_purchase$date_of_purchase,
                                           units = "days"))
```

```{r exploring data}
head(df_purchase)
summary(df_purchase)
```

```{r creating customer data for 2015}
# Load the library
library(sqldf)

# Create the data frame
cust_2015 <- sqldf("SELECT
                    cust_id, 
                    MIN(days_since) AS 'recency',
                    MAX(days_since) AS 'first_purchase',
                    COUNT(*) AS 'frequency',
                    AVG(purchase_amt) AS 'monvalue'
                  FROM
                    df_purchase
                  GROUP BY
                    cust_id")
```

```{r exploring 2015 customer data}
# Prerequisite exploring
head(cust_2015)
summary(cust_2015)

# Data distribution
hist(cust_2015$recency)
hist(cust_2015$first_purchase)
hist(cust_2015$frequency)
hist(cust_2015$monvalue, breaks = 100)
```

## Coding the Managerial Segmentation

This is the section where I should make a change. I will make a change based on what question want me to do. 

```{r generating segmentation}
# Create the segment variable
cust_2015$segment <- "NA"

# Create the segmentation
cust_2015$segment[which(cust_2015$recency > 365*3)] <- "inactive"
cust_2015$segment[which(cust_2015$recency <= 365*3 & cust_2015$recency > 365*2)] <- "cold"
cust_2015$segment[which(cust_2015$recency <= 365*2 & cust_2015$recency > 365*1)] <- "warm"
cust_2015$segment[which(cust_2015$recency <= 365)] = "active"
cust_2015$segment[which(cust_2015$segment == "warm" & cust_2015$first_purchase <= 365*2)] = "new warm"
cust_2015$segment[which(cust_2015$segment == "warm" & cust_2015$monvalue < 100)] <- "warm low value"
cust_2015$segment[which(cust_2015$segment == "warm" & cust_2015$monvalue >= 100)] <- "warm high value"
cust_2015$segment[which(cust_2015$segment == "active" & cust_2015$first_purchase <= 365)] = "new active"

# This is what different from previous exercise
cust_2015$segment[which(cust_2015$segment == "new active" & cust_2015$monvalue >= 100)] = "new active high"
cust_2015$segment[which(cust_2015$segment == "new active" & cust_2015$monvalue <= 100)] = "new active low"

cust_2015$segment[which(cust_2015$segment == "active" & cust_2015$monvalue < 100)] <- "active low value"
cust_2015$segment[which(cust_2015$segment == "active" & cust_2015$monvalue >= 100)] <- "active high value"

# Check the result
table(cust_2015$segment)
cust2015agg <- aggregate(x = cust_2015[, 2:5], by = list(cust_2015$segment), mean)
```

```{r ordering segmentation}
cust_2015$segment <- factor(x = cust_2015$segment, 
                           levels = c("inactive", "cold", "warm high value", "warm low value", "new warm", "active high value", "active low value", "new active high", "new active low"))

table(cust_2015$segment)
pie(table(cust_2015$segment))
aggregate(x = cust_2015[, 2:5], by = list(cust_2015$segment), mean)
```

I am done in creating the data for customers in 2015. Not forget to make modification for the customer 2014 data.

```{r creating 2014 customer data}
cust_2014 <- sqldf("SELECT
                    cust_id, 
                    MIN(days_since) - 365 AS 'recency',
                    MAX(days_since) - 365 AS 'first_purchase',
                    COUNT(*) AS 'frequency',
                    AVG(purchase_amt) AS 'monvalue'
                  FROM
                    df_purchase
                  WHERE
                    days_since > 365
                  GROUP BY
                    cust_id")
```

Let's segment the 2014 customer data using the same code as the 2015 customer data.

```{r generating segmentation for 2014 customer data}
# Reset the variable first
cust_2014$segment = "NA"

# Create the segmentation
cust_2014$segment[which(cust_2014$recency > 365*3)] = "inactive"
cust_2014$segment[which(cust_2014$recency <= 365*3 
                             & cust_2014$recency > 365*2)] = "cold"
cust_2014$segment[which(cust_2014$recency <= 365*2 
                             & cust_2014$recency > 365*1)] = "warm"
cust_2014$segment[which(cust_2014$recency <= 365)] = "active"
cust_2014$segment[which(cust_2014$segment == "warm" 
                             & cust_2014$first_purchase <= 365*2)] = "new warm"
cust_2014$segment[which(cust_2014$segment == "warm" 
                             & cust_2014$monvalue < 100)] = "warm low value"
cust_2014$segment[which(cust_2014$segment == "warm" 
                             & cust_2014$monvalue >= 100)] = "warm high value"
cust_2014$segment[which(cust_2014$segment == "active" 
                             & cust_2014$first_purchase <= 365)] = "new active"

cust_2014$segment[which(cust_2014$segment == "new active" 
                        & cust_2014$monvalue < 100)] <- "new active low"
cust_2014$segment[which(cust_2014$segment == "new active" 
                        & cust_2014$monvalue >= 100)] <- "new active high"

cust_2014$segment[which(cust_2014$segment == "active" 
                             & cust_2014$monvalue < 100)] = "active low value"
cust_2014$segment[which(cust_2014$segment == "active" 
                             & cust_2014$monvalue >= 100)] = "active high value"

# Check the result
table(cust_2014$segment)
pie(table(cust_2014$segment))
cust2014agg <- aggregate(x = cust_2014[, 2:5], by = list(cust_2014$segment), mean)
```

```{r ordering segmentation}
cust_2014$segment <- factor(x = cust_2014$segment, 
                           levels = c("inactive", "cold", "warm high value", "warm low value", "new warm", "active high value", "active low value", "new active high", "new active low"))

table(cust_2014$segment)
pie(table(cust_2014$segment))
aggregate(x = cust_2014[, 2:5], by = list(cust_2014$segment), mean)
```

## Calculating Revenue Generated From Customers in 2014 and 2015

```{r creating 2015 revenue data frame}
df_purchase$year_of_purchase <- format(df_purchase$date_of_purchase, format="%Y")

rev_2015 <- sqldf("SELECT
                    cust_id,
                    SUM(purchase_amt) AS 'rev_2015'
                  FROM
                    df_purchase
                  WHERE
                    year_of_purchase = '2015'
                  GROUP BY
                    cust_id")

summary(rev_2015)
```

```{r creating actual customer and revenue in 2015}
actual <- merge(cust_2015, rev_2015, all.x = TRUE)
actual$rev_2015[is.na(actual$rev_2015)] = 0

aggregate(x = actual$rev_2015, by = list(cust_2015$segment), FUN = mean)
```

```{r creating forward 2014 customer and revenue in 2015}
forward <- merge(cust_2014, rev_2015, all.x = TRUE)
forward$rev_2015[is.na(forward$rev_2015)] = 0
```

```{r aggregating}
# Show average revenue per customer and per segment
r <- aggregate(x = forward$rev_2015, by = list(cust_2014$segment), mean)
print(r)

# Re-order and display results
r = r[order(r$x, decreasing = TRUE), ]
print(r)
barplot(r$x, names.arg = r$Group.1)
```

## Questions : Solve the Problem

1. How many "new active low" customers were there in 2015? *Answer: 1249*

```{r q1}
table(cust_2015$segment)
```

2. The number of "new active high" customers has increased between 2014 and 2015. What is the rate of that increase? *Answer: 29.6%*

```{r q2}
cust_2015 %>% 
  count(segment) %>% 
  filter(segment == "new active high")

cust_2014 %>% 
  count(segment) %>% 
  filter(segment == "new active high")

(263 - 203) / 203
```

3. Regarding the customers  who belonged to the "new warm" segment in 2014, what was there expected revenue, all things considered, in 2015? *Answer: 5.064*

```{r q3}
r$x[which(r$Group.1 == "new warm")]
```

4. In terms of expected revenue, which segment groups the least profitable customers? *Answer: Inactive*

```{r q4}
r %>% 
  slice(which.min(x)) %>% 
  select(Group.1)
```

5. Looking at segment description, what is the average purchase amount of a customer who belongs to the "new active high" segment? *Answer: 283.38*

```{r q5}
aggregate(x = cust_2015[, "monvalue"], by = list(cust_2015$segment), mean)
```