---
title: "W4-1_BuildingAPredictiveModel"
author: "Taufiqur Rohman"
date: '2022-04-04'
output: html_document
---

**OBJECTIVE:** 
- Creating models for predicting active or inactive, and expected revenue for the 2015 customers from the 2014 customers data.

## Reading and Cleaning the Data

Read the data using read.delim, because the txt source file is separated by tab. And then, similar to previous module, I give name to every columns. Also, do not forget to convert the data format of date_of_purchase into date and calculate the recency of last purchase.

```{r read data frame}
df_purchase <- read.delim("purchases.txt", header = FALSE, sep = "\t", dec = ".")
colnames(df_purchase) <- c("cust_id", "purchase_amt", "date_of_purchase")

# Converting data format and adding Recency column
df_purchase$date_of_purchase <- as.Date(df_purchase$date_of_purchase, "%Y-%m-%d")
df_purchase$days_since <- as.numeric(difftime(time1 = "2016-01-01",
                                           time2 = df_purchase$date_of_purchase,
                                           units = "days"))
```

After that, I create the 2014 customer data first.

```{r creating 2014 customer data}
cust_2014 <- sqldf("SELECT
                    cust_id, 
                    MIN(days_since) - 365 AS 'recency',
                    MAX(days_since) - 365 AS 'first_purchase',
                    COUNT(*) AS 'frequency',
                    AVG(purchase_amt) AS 'monvalue',
                    MAX(purchase_amt) AS 'maxvalue'
                  FROM
                    df_purchase
                  WHERE
                    days_since > 365
                  GROUP BY
                    cust_id")
```

Then, I collect the 2015 revenue data as the dependent variable, or the target variable.

```{r create 2015 revenue data frame}
df_purchase$year_of_purchase <- format(df_purchase$date_of_purchase, format="%Y")

rev_2015 <- sqldf("SELECT
                    cust_id,
                    SUM(purchase_amt) AS 'rev_2015'
                  FROM
                    df_purchase
                  WHERE
                    year_of_purchase = '2015'
                  GROUP BY
                    cust_id")
```

## Creating the model

Now, let's create the model. First, I merge the customer 2014 data and 2015 revenue data to train the model. After that, I put the binary value for the active or inactive.

```{r building the model}
df_train <- merge(cust_2014, rev_2015, all.x = TRUE)
df_train$rev_2015[is.na(df_train$rev_2015)] = 0
df_train$active_2015 <- as.numeric(df_train$rev_2015 > 0) # Convert TRUE or FALSE to binary 1 or 0

# Checking the data
head(df_train)
summary(df_train)
```

Now the data is clean, let's build the model. First, to predict whether a customer will be active or inactive a year from now, I use the multinominal logistic regression.

```{r model for active or inactive}
library(nnet)
active_model <- multinom(formula = active_2015 ~ recency + first_purchase + frequency + monvalue + maxvalue, data = df_train)

coef <- summary(active_model)$coefficients
std.err  <- summary(active_model)$standard.errors

print(coef)
print(std.err)
print(coef / std.err) 

# Coef / std. err : Parameter for significance value. Below or above +2 / -2 is significance
```

After that, I create model to predict expected revenue a year from now using linear regression model.

```{r model for expected revenue} 
df_train_mon <- df_train %>%
  filter(active_2015 == 1)

head(df_train_mon)
summary(df_train_mon)

mon_model <- lm(rev_2015 ~ monvalue + maxvalue, data = df_train_mon)
summary(mon_model)
```

Let's test the accuracy of the model now.

```{r plot for the model accuracy}
plot(x = df_train_mon$rev_2015, y = mon_model$fitted.values)
```

I found that the data is very dispersed at the big value. Let's re calibrate the model to improve the accuracy by taking log into the predictors and target value.

```{r model for expected revenue (improved accuracy)}
mon_model <- lm(log(rev_2015) ~ log(monvalue) + log(maxvalue), data = df_train_mon)
summary(mon_model)

plot(x = log(df_train_mon$rev_2015), y = mon_model$fitted.values)
```

## Predict the 2015 Customer Data

Before predicting the 2015 customer data, I want to create the 2015 customer data frame first.

```{r}
cust_2015 <- sqldf("SELECT
                    cust_id, 
                    MIN(days_since) AS 'recency',
                    MAX(days_since) AS 'first_purchase',
                    COUNT(*) AS 'frequency',
                    AVG(purchase_amt) AS 'monvalue',
                    MAX(purchase_amt) AS 'maxvalue'
                  FROM
                    df_purchase
                  WHERE
                    days_since > 365
                  GROUP BY
                    cust_id")
```

Now, let's predict the expected revenue and idle probabilty of the customers using the multinominal logistic and linear regression that I have built before.

```{r}
cust_2015$prob_predict <- predict(object = active_model, newdata = cust_2015, type = "probs")
cust_2015$rev_predict <- exp(predict(object = mon_model, newdata = cust_2015))

# Create score prediction for every customer
cust_2015$score_predict <- cust_2015$prob_predict * cust_2015$rev_predict

summary(cust_2015$prob_predict)
summary(cust_2015$rev_predict)
summary(cust_2015$score_predict)
```

Looking from the summary, I know that the expected revenue from every customer one year from 2015 will be around 63.25 dollar. Also, there will be a customer that spend around 3800 dollar which is very high. The minimum a customer will spend is 6.54 dollar.