---
title: "W5-2_QuizModule4"
author: "Taufiqur Rohman"
date: '2022-04-06'
output: html_document
---

To validate Module 4 and correctly answer the questions will require that you perform the following exercise first: take the code in the file named module4.R, and modify it such as, in the simulations, from 2016 to 2025, 1,000 new customers are acquired every year and added to the database. Tip: it will require that you add a line of code between the lines #99 and #100 to specify that segment 8 has an influx of 1,000 new customers every year. 

## Reading and Cleaning Data Frame

First, I load the data first into the R environment.

```{r reading and cleaning data}
df_purchase <- read.delim("purchases.txt", header = FALSE, sep = "\t", dec = ".")
colnames(df_purchase) <- c("cust_id", "purchase_amt", "date_of_purchase")

# Converting data format and adding Recency column
df_purchase$date_of_purchase <- as.Date(df_purchase$date_of_purchase, "%Y-%m-%d")
df_purchase$days_since <- as.numeric(difftime(time1 = "2016-01-01",
                                           time2 = df_purchase$date_of_purchase,
                                           units = "days"))
```

After that, like the previous module, I create the 2015 and 2014 customer data.

```{r creating data frame for 2014 and 2015 customers}
# Load the library
library(sqldf)

# Create the 2015 customer data frame
cust_2015 <- sqldf("SELECT
                    cust_id, 
                    MIN(days_since) AS 'recency',
                    MAX(days_since) AS 'first_purchase',
                    COUNT(*) AS 'frequency',
                    AVG(purchase_amt) AS 'monvalue'
                  FROM
                    df_purchase
                  GROUP BY
                    cust_id")

# Create the 2014 customer data frame

cust_2014 <- sqldf("SELECT
                    cust_id, 
                    MIN(days_since) - 365 AS 'recency',
                    MAX(days_since) - 365 AS 'first_purchase',
                    COUNT(*) AS 'frequency',
                    AVG(purchase_amt) AS 'monvalue'
                  FROM
                    df_purchase
                  WHERE
                    days_since > 365
                  GROUP BY
                    cust_id")
```

After the data is loaded and data frame for 2015 and 2014 customer data is made, I made segmentation data frame for 2015 and 2014 customers data.

```{r generating segmentation}
# Reset the variable first
cust_2015$segment = "NA"
cust_2014$segment = "NA"

# Create the 2015 segmentation
cust_2015$segment[which(cust_2015$recency > 365*3)] = "inactive"
cust_2015$segment[which(cust_2015$recency <= 365*3 & cust_2015$recency > 365*2)] = "cold"
cust_2015$segment[which(cust_2015$recency <= 365*2 & cust_2015$recency > 365*1)] = "warm"
cust_2015$segment[which(cust_2015$recency <= 365)] = "active"
cust_2015$segment[which(cust_2015$segment == "warm" & cust_2015$first_purchase <= 365*2)] = "new warm"
cust_2015$segment[which(cust_2015$segment == "warm" & cust_2015$monvalue < 100)] = "warm low value"
cust_2015$segment[which(cust_2015$segment == "warm" & cust_2015$monvalue >= 100)] = "warm high value"
cust_2015$segment[which(cust_2015$segment == "active" & cust_2015$first_purchase <= 365)] = "new active"
cust_2015$segment[which(cust_2015$segment == "active" & cust_2015$monvalue < 100)] = "active low value"
cust_2015$segment[which(cust_2015$segment == "active" & cust_2015$monvalue >= 100)] = "active high value"

cust_2015$segment = factor(x = cust_2015$segment, 
                           levels = c("inactive", "cold", "warm high value", "warm low value", "new warm", "active high value", "active low value", "new active"))

# Create the 2014 segmentation
cust_2014$segment[which(cust_2014$recency > 365*3)] = "inactive"
cust_2014$segment[which(cust_2014$recency <= 365*3 & cust_2014$recency > 365*2)] = "cold"
cust_2014$segment[which(cust_2014$recency <= 365*2 & cust_2014$recency > 365*1)] = "warm"
cust_2014$segment[which(cust_2014$recency <= 365)] = "active"
cust_2014$segment[which(cust_2014$segment == "warm" & cust_2014$first_purchase <= 365*2)] = "new warm"
cust_2014$segment[which(cust_2014$segment == "warm" & cust_2014$monvalue < 100)] = "warm low value"
cust_2014$segment[which(cust_2014$segment == "warm" & cust_2014$monvalue >= 100)] = "warm high value"
cust_2014$segment[which(cust_2014$segment == "active" & cust_2014$first_purchase <= 365)] = "new active"
cust_2014$segment[which(cust_2014$segment == "active" & cust_2014$monvalue < 100)] = "active low value"
cust_2014$segment[which(cust_2014$segment == "active" & cust_2014$monvalue >= 100)] = "active high value"

cust_2014$segment = factor(x = cust_2014$segment, 
                           levels = c("inactive", "cold", "warm high value", "warm low value", "new warm", "active high value", "active low value", "new active"))

# Check the result
table(cust_2015$segment)
cust2015agg <- aggregate(x = cust_2015[, 2:5], by = list(cust_2015$segment), mean)

table(cust_2014$segment)
cust2014agg <- aggregate(x = cust_2014[, 2:5], by = list(cust_2014$segment), mean)
```

This is the important thing. I create the transition matrix first to know rate of consumer conversion from one segment to another.

```{r creating transition matrix}
df_trans <- merge(x = cust_2014, y = cust_2015, by = "cust_id", all.x = TRUE)
head(df_trans)
colnames(df_trans)

# Calculate the proportion
trans_matrix <- table(df_trans$segment.x, df_trans$segment.y)
print(trans_matrix)

# Convert it into percentage
trans_matrix <- trans_matrix / rowSums(trans_matrix)
print(trans_matrix)
```

The data is ready now and I can start calculating the expected customer lifetime value.

## Transforming Data for Expected CLV 

First, let's set up the matrix table for multiplier of the transition matrix.

```{r creating the matrix for the multiplier}
# Create a multiplier matrix first
# 8 rows for number of segments
# 11 columns for number of period we want to predict, until 2025
multiplier <- matrix(nrow = 8, ncol = 11)

# Inputting the present data to the matrix
multiplier[, 1] = table(cust_2015$segment)

# Giving column names, 2015 until 2025
colnames(multiplier) <- 2015 : 2025

# Giving row names by the segment
row.names(multiplier) <- levels(cust_2015$segment)
```

Here's the part where I should make a change. I have to add 1000 customers every year in the "new active segment"

```{r predicting customer lifetime value}
for (i in 2:11) {
   multiplier[, i] <- multiplier[, i-1] %*% trans_matrix
   multiplier[8, i] <- 1000
}
```

Let's explore some data inside of it. 

```{r creating the barplot}
barplot(multiplier["inactive", ], 
        main = "Expected Inactive Customers", xlab = "Years", ylab = "Number of Customers")
barplot(multiplier["cold", ], 
        main = "Expected Cold Customers", xlab = "Years", ylab = "Number of Customers")
```

## Computing Customer Lifetime Value

Let's now calculate the customer lifetime value, which is the main objective of this analysis. First, let's put the revenue data in the vector and calculate revenue per segment, revenue per year, and cumulated revenue

```{r computing expected revenue generated by customer}
# Revenue data per segment is readily available, just put it into vector
rev_annual <- c(0, 0, 0, 0, 0, 323.57, 52.31, 79.17)

# Compute revenue per segment
rev_segment <- rev_annual * multiplier
print(rev_segment)

# Revenue per year
rev_annual <- colSums(rev_segment)
print(round(rev_annual))
barplot(rev_annual)

# Cumulated revenue
rev_cumulated <- cumsum(rev_annual)
print(rev_cumulated)
barplot(rev_cumulated)
```

**REMEMBER** that the future value of the money will be decreasing from time to time because of the inflation. Let's calculate the discount factor with inflation rate at 10% a year.

```{r create discount factor}
disc_rate <- 0.10
disc = 1 / ((1 + disc_rate) ^ ((1:11) - 1))
print(disc)
```

Now, multiply the discount factor with the revenue data that I have calculated earlier.

```{r computing discounted revenue}
# Discounted revenue per year
disc_rev_annual <- rev_annual * disc
print(disc_rev_annual)
barplot(disc_rev_annual)

# Discounted cumulated revenue per year
disc_rev_cumulated <- cumsum(disc_rev_annual)
print(disc_rev_cumulated)
barplot(disc_rev_cumulated)

# Comparing discounted vs. undiscounted revenue per year
barplot(disc_rev_annual)
lines(rev_annual)
```

For the conclusion, let's calculate how much money we can expect from current customers 10 years from now.

```{r expected total revenue}
disc_rev_cumulated[11] - disc_rev_cumulated[1]
```

We can expect that in 10 years from now, the company will get 2,150,640 dollar from the current customers.

## Questions : Problem to Solve

1. How many "inactive" customers does the model predict there will be in 2025? 

```{r}
round(multiplier["inactive", "2025"])

# Answer: 18,275 customers
```

2. What would the database be worth by 2025 (cumulated revenues, discounted) if a constant influx of 1,000 new customers was added every year till then? 

```{r}
disc_rev_cumulated[11] - disc_rev_cumulated[1]

# Answer: $3,308,873
```

3. Looking at the transition matrix, what is the likelihood of a "new warm" customer to become "cold"? 

```{r}
trans_matrix["new warm", "cold"]

# Answer: 91%
```

4. Looking at the transition matrix, the likelihood of transitioning from "new active" to "cold" customer is 0%. Why so? 

```{r}
trans_matrix["new active", "cold"]

# Answer: 
```

5. After the modification to the code, how many customers will there be in the database by 2022?

```{r}
sum(multiplier[, "2022"])

# Answer: 25,417 customers
```

